### 2013-09-27 structured_prediction/paper.tex
some moving around of figures and tables
k-means not capitalizes, minor fixes

In recent years there has been a wealth of research in methods for learning
structured prediction, as well as in their application in areas such as natural
language processing and computer vision (see \citet{nowozin2011structured}
for an introduction and \citet{blake2011markov} for a recent survey).
%
In this chapter, we first introduce the concepts and algorithms used in
structured prediction, in particular in maximum margin methods. We 
accompany this by a description of our open source implementation of these
algorithms, \pystruct.

@@

Structured prediction can be defined as making a prediction $f(x)$ by
maximizing a compatibility function $g$ between an input $x$ and possible
labels
$y$~\citep{nowozin2011structured}:
\begin{equation}
    f(x) = \argmax_{y \in \mathcal{Y}} g(x, y)
\end{equation}
Finding $y$ in the above equation is often referred to as inference
or prediction.
Most current approaches use linear functions, with some notable exceptions:
\citet{peng2009conditional} introduced a conditional random field for sequence
prediction where single-node potentials are computed using neural networks,
while \citet{nowozin2011decision} proposed a conditional random field for image
processing, where all potentials are computed using decision trees.

The linear approach leads to
\begin{equation}\eqlabel{main_equation}
    f(x) = \argmax_{y \in \mathcal{Y}}\  \theta^T \Phi(x, y).
\end{equation}
Here, $y$ is a \emph{structured} label, $\Phi$ is a joint feature function of
$x$ and $y$, and $\theta$ contains the parameters of the model. \emph{Structured} means
that $y$ is more complicated than a single output class, for example a label
for each word in a sentence or a label for each pixel in an image.
Learning structured prediction means learning the parameters $\theta$ from
training data.  The particular model that is used is completely encoded in
$\Phi(x, y)$, which manifests the relation between input $x$ and output $y$. As
$\mathcal{Y}$ is typically very large, it is crucial to exploit the particular
form of $\Phi(x, y)$ to solve the prediction problem of \Eqref{main_equation}.

@@

Arguably the most straight-forward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, it is natural
to work in a stochastic setting (see \citet{ratliff2007online}).\pagebreak\\
Given a model through $\Phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align*}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \Phi(x\hoch{i}, \hat{y}) - \Phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\numberthis\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \Phi(x\hoch{i}, y)
\end{align*}
The most commonly used update has the simple form
\begin{equation}
    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\Phi(x\hoch{i}, \hat{y}) - \Phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of step sizes.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
In practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopt the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
\citet{shalev2011pegasos} showed that this schedule achieves a convergence rate of $O(\frac{\ln T }{T})$.
\citet{lacoste2012block} and \citet{shamir2012stochastic} recently showed independently that
a rate of $O(\frac{1}{T})$ can be achieved using a novel averaging scheme:
\begin{equation}
    \bar{\theta}_{T} = \frac{2}{(T+1)(T+2)} \sum_{t=0}^T(t+1) \theta_t.
\end{equation}
This $t$-weighted averaging can be computed on-the-fly as
\begin{equation}\eqlabel{averaging}
    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}.
\end{equation}

@@

Very recently, two groups independently derived a very performant new algorithm.
\citet{shalev2012proximal} investigated stochastic dual coordinate descent applied
to structured support vector machines. \citet{lacoste2012block} started from the Frank-Wolfe algorithm %TODO details
and derived a block-coordinate version, where each block corresponds to the
constraints associated with a single training example.  Both groups showed that
closed-form line search is possible, and, surprisingly ended up with the exact
same algorithm and closely related convergence guarantees.  The algorithm
processes a single training example at a time, and can be applied while
remaining completely in the primal domain, making it applicable to very large
datasets.  We refer to the algorithm as SZLJSP, after the authors of the
two papers.
%
The SZLJSP algorithm has the same theoretical convergence rate of $O(\frac{1}{T})$
as stochastic subgradient decent, but has two distinguishing advantages:
\begin{description}
\item[Stopping Criterion] Both views of SZLJSP give rise to a dual objective,
    which can be used as a theoretically sound stopping criterion.
\item[Learning Rate] While there are several theoretical results on choosing
    the step size in stochastic subgradient descent, choosing a concrete
    schedule is often problematic in practice. By using analytic line-search,
    SZLJSP removes the need for any step size parameter, making application
    of the algorithm much simpler.
\end{description}

In practice, the SZLJSP algorithm is usually used with the weighted averaging
described in \Eqref{averaging}, as this is known to improve performance in the
closely related stochastic subgradient algorithms, and empirically improves
convergence.
\enlargethispage{10mm}
The detailed procedure as given by \citet{lacoste2012block} is shown in Algorithm~\ref{alg_szljsp}.

\section{Conditional Random Fields for Semantic Segmentation}\seclabel{semantic_segmentation}
%
\subsection{Fundamentals of Conditional Random Fields}
Structured models based on factor graphs, as described in \Secref{factor_graphs} are often
referred to as Conditional Random Fields (CRFs), pointing to their
probabilistic interpretation.
CRFs have been established as an important tool in many areas of computer
vision. Applications include dense stereo, optical flow, inpainting,
denoising, image editing, low-level segmentation and semantic segmentation.

@@

Using superpixels instead of pixels has large computational advantages.  A
model that uses superpixels contains much fewer variables than one that works
on pixel-level. Typical sizes for superpixels are in the hundreds of pixels,
leading to a hundred-fold decrease in the number of variables. There are also
semantic advantages. Using superpixels can constitute a form of regularization,
including prior knowledge about the structure of images. Modeling interactions
between superpixels also allows more distant parts of an image to interact,
instead of modeling long-term relations only indirectly over neighbors.
Finally, modeling decisions on a superpixel-level allows for more context for a
local decision, while semantic decisions on pixel-level are not meaningful near
object boundaries.  The main disadvantage of superpixel-based approaches is
that the initial over-segmentation can not be influenced by later reasoning.
This can be a problem if fine structures need to be segmented for which local
evidence does not suffice.
%
\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_org}&%
    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp}\\
    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_graph}&%
    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_gt}
\end{tabu}
\caption{%
    Schematic of CRFs over superpixels. From left to right: input image, superpixel segmentation
    using SLIC into about 100 superpixels, pairwise potential in the CRF, and desired labeling.
\figlabel{superpixel_illustration}}
\end{figure}

@@

% submodular/non-submodular? higher order?
\subsection{Previous Work}
Conditional random fields were first used in the context of semantic
segmentation by \citet{he2004multiscale}.  Who used contrastive divergence for
learning parameters. Many other early models, such as the one of
\citet{shotton2006textonboost}, did not learn parameters at all, but used
\emph{contrast-sensitive Potts potential}. These potentials have a smoothing
effect over labels by encouraging neighboring variables to take the same value.
The penalty for taking different values is dependent on color contrast between
the pixels.

@@

\subsection{Data, Features and Superpixels}
When evaluating learning algorithms in Chapter~\ref{ch:comparison} and
Chapter~\ref{ch:exact_learning} we use the well-established \textsc{Pascal} VOC 2010 and
MSRC-21 datasets, shown in the introduction in \Figref{pascal_msrc}.
As we focus on the learning algorithms in these chapters, we use features for
unary potentials from the literature.

We use the SLIC~\citep{achanta2012slic} algorithm to create superpixel for all
our experiments.  It has been shown to provide competitive results with a
minimum of computational complexity.  Algorithmically, SLIC simply computes a
$k$-means clustering over pixels. Each pixels is represented as 5D point, using
three color channels and the $x$ and $y$ coordinates in the image.  In a
post-processing step, small segments are removed. To make clustering of so many
points feasible, the search for the nearest cluster in $k$-means is restricted
to a local neighborhood in the image. An example of superpixels computed with
the SLIC algorithm can be found in \Figref{superpixel_illustration}.

The features and superpixel algorithm used for the RGB-D dataset NYU used in
Chapter~\ref{ch:nyu} will be discussed there.

>>>>

In recent years there has been a wealth of research in methods for learning
structured prediction, as well as in their application in areas such as natural
language processing and computer vision (see \citet{nowozin2011structured}
for an introduction and \citet{blake2011markov} for a recent survey).
%
In this chapter, we first introduce the concepts and algorithms used in
structured prediction, in particular in maximum margin methods. Then, we review
the use of CRFs in computer vision, and introduce our methods. Finally we give
a description of our open source implementation of structured learning
algorithms, \pystruct.

@@

Structured prediction can be defined as making a prediction $f(x)$ by
maximizing a compatibility function $g$ between an input $x$ and possible
labels
$y$~\citep{nowozin2011structured}:
\begin{equation}
    f(x) = \argmax_{y \in \mathcal{Y}} g(x, y)
\end{equation}
Finding $y$ in the above equation is often referred to as inference or
prediction.  We will use the most common approach of using a linear parametrization of $g$,
which leads to
\begin{equation}\eqlabel{main_equation}
    f(x) = \argmax_{y \in \mathcal{Y}}\  \theta^T \Phi(x, y).
\end{equation}
Here, $y$ is a \emph{structured} label, $\Phi$ is a joint feature function of
$x$ and $y$, and $\theta$ contains the parameters of the model. \emph{Structured} means
that $y$ is more complicated than a single output class, for example a label
for each word in a sentence or a label for each pixel in an image.
Learning structured prediction means learning the parameters $\theta$ from
training data.  The particular model that is used is completely encoded in
$\Phi(x, y)$, which manifests the relation between input $x$ and output $y$. As
$\mathcal{Y}$ is typically very large, it is crucial to exploit the particular
form of $\Phi(x, y)$ to solve the prediction problem of \Eqref{main_equation}.

@@

Arguably the most straight-forward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, it is natural
to work in a stochastic setting (see \citet{ratliff2007online}).\pagebreak\\
Given a model through $\Phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align*}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \Phi(x\hoch{i}, \hat{y}) - \Phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\numberthis\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \Phi(x\hoch{i}, y)
\end{align*}
The most commonly used update has the simple form
\begin{equation}
    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\Phi(x\hoch{i}, \hat{y}) - \Phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Here $\eta_t$ is a sequence of step sizes.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
In practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopt the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
\citet{shalev2011pegasos} showed that this schedule achieves a convergence rate of $O(\frac{\ln T }{T})$.
\citet{lacoste2012block} and \citet{shamir2012stochastic} recently showed independently that
a rate of $O(\frac{1}{T})$ can be achieved using a novel averaging scheme:
\begin{equation}
    \bar{\theta}_{T} = \frac{2}{(T+1)(T+2)} \sum_{t=0}^T(t+1) \theta_t.
\end{equation}
This $t$-weighted averaging can be computed on-the-fly as
\begin{equation}\eqlabel{averaging}
    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}.
\end{equation}

@@

Very recently, two groups independently derived a very performant new algorithm.
\citet{shalev2012proximal} investigated stochastic dual coordinate descent applied
to structured support vector machines. \citet{lacoste2012block} started from
the Frank-Wolfe algorithm and derived a block-coordinate version, where each
block corresponds to the constraints associated with a single training example.
Both groups showed that closed-form line search is possible, and, surprisingly
ended up with the exact same algorithm and closely related convergence
guarantees.  The algorithm processes a single training example at a time, and
can be applied while remaining completely in the primal domain, making it
applicable to very large datasets.  We refer to the algorithm as SZLJSP, after
the authors of the two papers.
%
The SZLJSP algorithm has the same theoretical convergence rate of $O(\frac{1}{T})$
as stochastic subgradient decent, but has two distinguishing advantages:
\begin{description}
\item[Stopping Criterion] Both views of SZLJSP give rise to a dual objective,
    which can be used as a theoretically sound stopping criterion.
\item[Learning Rate] While there are several theoretical results on choosing
    the step size in stochastic subgradient descent, choosing a concrete
    schedule is often problematic in practice. By using analytic line-search,
    SZLJSP removes the need for any step size parameter, making application
    of the algorithm much simpler.
\end{description}
The only disadvantage of SZLJSP compared to SSGD is the larger memory requirement.
While SSGD only needs to store a single copy of the parameters $\theta$, SZLJSP
needs to store a separate (though possibly sparse) copy for each training sample.

In practice, the SZLJSP algorithm is commonly used with the weighted averaging
described in \Eqref{averaging}, as this is known to improve performance in the
closely related stochastic subgradient algorithms, and empirically improves
convergence.
\enlargethispage{10mm}
The detailed procedure as given by \citet{lacoste2012block} is shown in Algorithm~\ref{alg_szljsp}.

\section[Conditional Random Fields for Semantic Segmentation]{Conditional Random Fields\linebreak for Semantic Segmentation}\seclabel{semantic_segmentation}
%
\subsection{Fundamentals of Conditional Random Fields}
Structured models based on factor graphs, as described in \Secref{factor_graphs} are often
referred to as Conditional Random Fields (CRFs), pointing to their
probabilistic interpretation.
CRFs have been established as an important tool in many areas of computer
vision. Applications include dense stereo, optical flow, inpainting,
denoising, image editing, low-level segmentation and semantic segmentation.

@@

Using superpixels instead of pixels has large computational advantages.  A
model that uses superpixels contains much fewer variables than one that works
on pixel-level. Typical sizes for superpixels are in the hundreds of pixels,
leading to a hundred-fold decrease in the number of variables. There are also
semantic advantages. Using superpixels can constitute a form of regularization,
including prior knowledge about the structure of images. Modeling interactions
between superpixels also allows more distant parts of an image to interact,
instead of modeling long-term relations only indirectly over neighbors.
Finally, modeling decisions on a superpixel-level allows for more context for a
local decision, while semantic decisions on pixel-level are not meaningful near
object boundaries.  The main disadvantage of superpixel-based approaches is
that the initial over-segmentation can not be influenced by later reasoning.
This can be a problem if fine structures need to be segmented for which local
evidence does not suffice.
%
\begin{figure}
    \begin{tabu} to \linewidth{@{}XX@{}}
    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_org}&%
    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp}\\
    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_graph}&%
    \includegraphics[width=\linewidth]{structured_prediction/images/scene_sp_gt}
\end{tabu}
\caption{%
    Schematic of CRFs over superpixels. From left to right: input image, superpixel segmentation
    using SLIC into about 100 superpixels, pairwise potential in the CRF, and desired labeling of superpixels.
\figlabel{superpixel_illustration}}
\end{figure}

@@

% submodular/non-submodular? higher order?
\subsection{Data, Features and Superpixels}
When evaluating learning algorithms in Chapter~\ref{ch:comparison} and
Chapter~\ref{ch:exact_learning} we use the well-established \textsc{Pascal} VOC 2010 and
MSRC-21 datasets, shown in the introduction in \Figref{pascal_msrc}.
As we focus on the learning algorithms in these chapters, we use features for
unary potentials from the literature.

We use the SLIC~\citep{achanta2012slic} algorithm to create superpixel for all
our experiments.  It has been shown to provide competitive results with a
minimum of computational complexity.  Algorithmically, SLIC simply computes a
$k$-means clustering over pixels. Each pixels is represented as 5D point, using
three color channels and the $x$ and $y$ coordinates in the image.  In a
post-processing step, small segments are removed. To make clustering of so many
points feasible, the search for the nearest cluster in $k$-means is restricted
to a local neighborhood in the image. An example of superpixels computed with
the SLIC algorithm can be found in \Figref{superpixel_illustration}.
%
The features and superpixel algorithm used for the RGB-D dataset NYU used in
Chapter~\ref{ch:nyu} will be discussed there.

\subsection{Previous Work}
Conditional random fields were first used in the context of semantic
segmentation by \citet{he2004multiscale}.  Who used contrastive divergence for
learning parameters. Many other early models, such as the one of
\citet{shotton2006textonboost}, did not learn parameters at all, but used
\emph{contrast-sensitive Potts potential}. These potentials have a smoothing
effect over labels by encouraging neighboring variables to take the same value.
The penalty for taking different values is dependent on color contrast between
the pixels.

@@
